\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,latexsym,amssymb,verbatim}
\usepackage[margin=1in]{geometry}
\pagestyle{empty}
\begin{document}

\begin{center}
{\bf \noindent {\Large J}OURNAL OF {\Large I}NQUIRY-{\Large B}ASED
{\Large L}EARNING
\medskip
\noindent IN {\Large M}ATHEMATICS}
\end{center}

\medskip
\begin{center}
{\large \bf Course Notes Submission Form}
\end{center}

Please email this form to the Course Notes Submissions Editor with any course notes submission. This information will be made available to JIBLM users when your course notes are published.

\begin{enumerate}

\item Author Name(s): David J. Hunter

\item University: Westmont College

\item Mailing Address: 955 La Paz Road, Santa Barbara, CA, 93108

\item Email Address: \texttt{dhunter@westmont.edu}

\item Title of Course Notes: \textit{Abstract Algebra I \& II}

\item Abstract: These notes provide an introduction to abstract algebra for advanced undergraduate mathematics majors. There should be enough material for a two-semester sequence. Chapters 1--9 constitute the first semester of this sequence, which covers groups, rings, and fields. Chapters 10--13 are intended for a second-semester course for more advanced students, presenting some advanced group theory, including Sylow Theory, with the main focus of introducing students to Galois Theory.

\item History of Use: (JIBLM publishes only classroom-tested course notes. Please list the history of the use of these course notes, by semester, including the names of the institutions where it was used.)

The original version of Chapters 1--5 of these notes were written by Margaret Morrow, and appeared as Issue 26 of JIBLM (May 2012). I have revised these chapters, and have added Chapters 6--13 to expand the notes into a two-semester sequence. I have included the designation ``Hunter $<$ Morrow'' on the title page.

I have taught through both semesters of these notes twice:

\begin{itemize}
    \item MA-110, Modern Algebra, Westmont College, Fall 2015.
    \item MA-111, Advanced Modern Algebra, Westmont College, Spring 2016.
    \item MA-110, Modern Algebra, Westmont College, Fall 2017.
    \item MA-111, Advanced Modern Algebra, Westmont College, Spring 2018.
\end{itemize}
\end{enumerate}

\end{document}
