---
geometry: left=1.0in, right=1.0in, top=0.8in, bottom=0.4in
classoption: twoside
output: 
  pdf_document:
    keep_tex: true
    includes:
      in_header: latexheader.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
```

# Advanced Abstract Algebra (MA-111) Westmont College, Spring 2024

## What is this course about?

Here's what the catalog says: "(Four credit hours) Prerequisite: MA-110. In 1832, when \text{\'{E}variste} Galois was 20 years old, he was shot and killed in a duel. The night before, he wrote a letter to a colleague detailing his work on some of the most profound mathematical questions of the time. This course investigates the work of Galois, combining advanced group theory and field theory to understand the nature of polynomial equations. (Additional topics may be included at the discretion of the instructor.)" Here's a list of the topics we will cover:
  \begin{itemize}[itemsep=0pt, parsep=1pt, partopsep=-5pt]
      \item \textbf{Advanced Group Theory:}  Symmetry, group actions, Sylow theory.
      \item \textbf{Field Theory:} Fields and extensions, algebraic extensions.
      \item \textbf{Galois Theory:} Isomorphisms of fields, splitting fields, separability and perfect fields, the Galois group, the fixed field, the Fundamental Theorem, solvability by radicals.
     \item \textbf{Groups in Topology:} (time permitting)
  \end{itemize}


## How is the course structured?

Resuming our work from last semester, we will work through a series of problems that will complete a comprehensive coverage of two semesters of advanced undergraduate or beginning graduate algebra. Ideally, you will supply the answers. This semester we will use a Wiki to build a collaborative set of solutions to the problem sequence.

\begin{description}
\item[Prework:] On most days, there will be a written assignment due the night before class. My expectation is that these prework assignments will be hand written, and the solutions need not be perfect. Use a scanner app (e.g., Genius Scan, Adobe Scan) to make a PDF of your prework and submit it on Canvas. If your prework spans multiple pages, please combine all pages into a single PDF file. In general, you will receive full credit for less-than-perfect prework assignments, as long as you make some progress on each problem.  Making progress can include writing down the first and last lines of the proof, and formulating some questions about what you didn't understand. I won't be making significant corrections to the prework.
\item[Assigned Presenters:] Each student will be assigned to be the   \textbf{presenter} for a given problem each day. Try to have a mostly complete and correct solution ready for the problem that you are presenting. If you need help, stop by student hours. Most of our class time will be spent presenting and discussing these solutions.
\item[Assigned Wiki Authors:]  For each problem that is presented, another student (distinct from the presenter) will be assigned to write up a final draft of the solution of the class Wiki.
\end{description}

The Presenter and Wiki Author assignments will appear on the Wiki: https://github.com/djhunter/ma111s24/wiki

## Is there a textbook?

The skeleton of the text for this course is a revised and expanded version of \textit{Introductory Abstract Algebra}, by Margaret L. Morrow, \textit{Journal of Inquiry Based Learning in Mathematics} \textbf{26}, (2012). Morrow's notes provide a gentle, inquiry-based introduction to group theory. I have added several chapters to these notes covering rings and fields. By the end of the semester, our class Wiki should contain complete solutions to all the problems in Chapters 10--13.

## How are grades determined?

```{r echo=FALSE}
kbl(data.frame(x = c("Prework Rough Drafts",  "Presentations and Participation", "Wiki Writeups", "Exams (3)", "Final Exam"),
                 y = c("10%", "15%", "15%", "15% each", "15%")), col.names = NULL, booktabs = TRUE)
```

## What other policies should students be aware of?

If you miss a significant number of classes, you will almost definitely do poorly in this class.  If you miss more than six classes without a valid excuse, I reserve the right to terminate you from the course with a failing grade.  Work missed (including tests) without a valid excuse will receive a zero.

I expect you to check your email on a regular basis.  If you use a non-Westmont email account, please forward your Westmont email to your preferred account. I'll send out notices on Canvas, so make sure you receive Canvas notifications in your email.

Learning communities function best when students have academic integrity.  Cheating is primarily an offense against your classmates because it undermines our learning community.  Therefore, dishonesty of any kind may result in loss of credit for the work involved and the filing of a report with the Provost's Office. Major or repeated infractions may result in dismissal from the course with a failing grade. Be familiar with the College's plagiarism policy, found at https://www.westmont.edu/office-provost/academic-program/academic-integrity-policy.

In particular, providing someone with an electronic copy of your work is a breach of the academic integrity policy. Do not email, post online, or otherwise disseminate any of the work that you do in this class. If you keep your work on a repository, make sure it is private. You may work with others on the assignments, but make sure that you write or type up your own answers yourself. You are on your honor that the work you hand in represents your own understanding.

## Other Information

\begin{description} 

\item[Professor:] David J. Hunter, Ph.D.
  (\verb!dhunter@westmont.edu!). Student hours are on Tuesdays from 1:00--3:20pm and Thursdays from 2:00--4:40pm in Winter Hall 303.

\item[Accommodations for Students with Disabilities:] Students who choose to disclose a disability, diagnosis, or injury are encouraged to contact the Office of Disability Services (ODS) as early as possible in the semester to discuss potential accommodations for this course. Formal accommodations will only be granted for students whose diagnoses have been verified by the ODS. Accommodations are designed to minimize the impact of a diagnosis and to ensure equal access to programs for all students who have a diagnosed condition that impacts their participation in college activities. Please contact \href{mailto:ods@westmont.edu}{\tt ods@westmont.edu} or visit the website for more information: \url{https://www.westmont.edu/disability-services-welcome}. Seth Miller, Director of ODS and the ODS team are located upstairs in Voskuyl Library 310, 311A. 

\item[Program and Institutional Learning Outcomes:] The
         mathematics department at Westmont College has formulated the
         following learning outcomes for all of its classes. (PLO's)
\begin{enumerate}[noitemsep]
\item Core Knowledge: Students will demonstrate knowledge of the
                  main concepts, skills, and facts of the discipline of
                  mathematics.
\item Communication: Students will be able to communicate mathematical ideas
     following the standard conventions of writing or speaking in the
     discipline.
\item Creativity: Students will demonstrate the ability to formulate and make
     progress toward solving non-routine problems.
\item Christian Connection: Students will incorporate their mathematical skills
     and knowledge into their thinking about their vocations as followers of
     Christ.
         \end{enumerate}
         In addition, the faculty of Westmont College have established common
         learning outcomes for all courses at the institution
         (ILO's). These outcomes are summarized as follows:
(1) Christian Understanding, Practices, and Affections,
(2) Global Awareness and Diversity,
(3) Critical Thinking,
(4) Quantitative Literacy,
(5) Written Communication,
(6) Oral Communication, and
(7) Information Literacy.

\item[Course Learning Outcomes:] The above outcomes are reflected in the
     particular learning outcomes for this course.
     After taking this course, you should be able
     to:
    \begin{itemize}
        \item Demonstrate understanding of modern abstract algebra.
             (PLO 1, ILOs 3,4)
        \item Write mathematical arguments according to the
             standards of the discipline. (PLO 2,
              ILOs 3,5)
        \item Construct solutions to novel problems,
               demonstrating perseverance in the face of open-ended or
               partially-defined contexts. (PLO 3, ILO 3)
        \item Consider the theological implications of the subject matter. (PLO 4, ILO 1)
    \end{itemize}
These outcomes will be assessed by written prework assignments, typeset final drafts, and written exams, as described above.

\end{description}
